---
title: "HOLC"
author: "Tong_Li"
date: "2021/12/11"
output:
  html_document: default
  pdf_document: default
---

## imput package

```{r setup, include=FALSE}
library(tidyverse)
library(tidyr)
library(sf)
library(here)
library(fs)
library(stringr)
library(utils)
library(sp)
library(cowplot)#draw scatter plots
```

## read shpfiles
```{r}
#census的文件名我改了，原始的叫geo_export_65990e0b-9fcf-4f65-be7a-573f665275e3.shp
census_shape <- st_read(here::here("Data", "Los_Angeles_census_tracts.shp"))%>%
  st_transform(., crs=32611)


#export csv file to check NA values 给我自己用的，可以不管
#write.csv(HOLC,"C:\\Users\\nekopi\\Desktop\\TEST1.csv", row.names = FALSE)

HOLC <- st_read(here::here("Data","cartodb-query.shp"))%>%
    st_transform(., crs=32611)


# 查找NA值
which(is.na(HOLC), arr.ind=TRUE)

#check out coordinate
#crs(Bands_need)
#crs(census_shape)
#crs(HOLC)
```
```{r}
# List your raster files excluding band 8 using the patter argument
Bands_need<-dir_info(here::here("Data"))%>%
  dplyr::filter(str_detect(path, "[B456710].TIF")) %>%
  #dplyr::filter(str_detect(path, "B10.TIF"))%>%
  dplyr::filter(str_detect(path, "B1.TIF", negate=TRUE))%>%
  dplyr::filter(str_detect(path, "B11", negate=TRUE))%>%
  dplyr::select(path)%>%
  arrange()%>%
  pull()%>%
  as.character()%>%
  # Load our raster layers into a stack
  raster::stack()

#only B10.tiff
Thermal<-dir_info(here::here("Data"))%>%
  dplyr::filter(str_detect(path, "B10.TIF")) %>%
  dplyr::select(path)%>%
  arrange()%>%
  pull()%>%
  as.character()%>%
  # Load our raster layers into a stack
  raster::stack()

thermal_crop <- Thermal %>%
    raster::crop(.,HOLC)

thermal_mask <- thermal_crop %>%
    raster::mask(.,HOLC)
    raster::plot(thermal_mask)

#用全部的 rasters堆叠并裁剪
lsatmask1 <- Bands_need %>%
  # now crop our temp data to the extent
  raster::crop(.,HOLC)
  raster::plot(lsatmask1)

lsat2 <- lsatmask1%>%
    raster::mask(.,  HOLC)
    raster::plot(lsat2)
    
# rename rasters in bands
names(Bands_need) <- c('thermal','red', 'NIR', 'SWIR1', 'SWIR2') 

#check out bands info
Bands_need

```

```{r}
#crop temp data to extent
censuscrop <- Bands_need %>%
  raster::crop(.,census_shape)
  raster::plot(censuscrop)
  
censusmask <- censuscrop %>%
  raster::mask(.,census_shape)
  raster::plot(censusmask)
```

make a function called NDVIfun
```{r}
NDVIfun <- function(NIR, Red) {
  NDVI <- (NIR - Red) / (NIR + Red)
  return(NDVI)
}

# 调用NDVI function
ndvi <- NDVIfun(lsat2$NIR, lsat2$red)

ndvi %>%
  plot(.,col = rev(terrain.colors(10)), main = "Landsat-NDVI")
```

NDVI
```{r}
#NDVI <- (lsat2$NIR - lsat2$red)/ (lsat2$NIR + lsat2$red)

# you need the sp package loaded for this...

#NDVI %>%
  #plot(.,col = rev(terrain.colors(10)), main = "Landsat-NDVI")
```
计算参数
```{r}
library(RStoolbox)

TOA <- (0.0003342 *
  thermal_mask) + 0.1

K1 <- as.numeric(774.8853)

K2 <- as.numeric(1321.0789)
  
Brighttemp <- (K2 / log((K1 / TOA) + 1))


facveg <- (ndvi-0.2/0.5-0.2)^2

emiss <- (0.004*facveg)+0.986
```

```{r}
Boltzmann <- 1.38*10e-23
Plank <- 6.626*10e-34
c <- 2.998*10e8

p <- Plank*(c/Boltzmann)

lambda <- 1.09e-5


#LST是下面还会用到的有温度数据的raster
LST <- Brighttemp/(1 +(lambda*Brighttemp/p)*log(emiss))


LST <- LST-273.15
raster::plot(LST)


```
NDBI
```{r}
NDBI=((lsat2$SWIR1-
         lsat2$NIR)/
        (lsat2$SWIR1+
        lsat2$NIR))

NDBI %>%
  raster::writeRaster(., "NDBI", 
              bylayer=TRUE, 
              format='GTiff', 
              overwrite=TRUE)

NDBI %>%
  plot(.,col = rev(terrain.colors(10)), main = "Landsat-NDBI")
```
```{r}

#这个chunk是为了筛选出graded的区域，level，面积

census_shape$area_of_extract <- census_shape%>%st_area()
pi <- st_intersection(HOLC, census_shape)

pi$intersect_area <- pi %>%
  st_area()

pitestA <- pi %>%
  group_by(ct10) %>%
  filter(holc_grade== "A") %>%
  summarise(A_area=sum(intersect_area))

pitestB <- pi %>%
  group_by(ct10) %>%
  filter(holc_grade== "B") %>%
  summarise(B_area=sum(intersect_area))

pitestC <- pi %>%
  group_by(ct10) %>%
  filter(holc_grade== "C") %>%
  summarise(C_area=sum(intersect_area))

pitestD <- pi %>%
  group_by(ct10) %>%
  filter(holc_grade== "D") %>%
  summarise(D_area=sum(intersect_area))

```


HOCL in census tracts
```{r}
#用HOLC的范围裁切census tract
census_sub <- census_shape[HOLC,, op=st_intersects]
#计算每个census的面积
census_sub <- census_sub %>%
  dplyr::mutate(census_area=(st_area(.)))
#计算每个裁切过的census的面积
intersect <- st_intersection(HOLC, census_sub)%>%
    dplyr::mutate(intersect_area=(st_area(.)))

tm_shape(intersect) +
  tm_polygons(col = "blue")+
  tm_shape(HOLC) +
  tm_borders(col="black")

```
deal with rasters
```{r}
#从raster堆栈中提取区域和温度信息
LST_per_census_within_HOLC <- raster::extract(LST, census_sub, fun=mean, na.rm=TRUE, df=TRUE)

#和shp对应
LST_per_census_within_HOLC<-LST_per_census_within_HOLC %>%
  mutate(ID=census_sub$ct10)%>%
  rename(temp=layer.1)
```


```{r}
# 把提取出来的A~D等级的各个区域再join_back回census

census_shapeA <- census_shape %>%
  st_drop_geometry()%>%
  left_join(., 
            pitestA,
            by = c("ct10" = "ct10"))

census_shapeAB <- census_shapeA %>%
  left_join(., 
            pitestB,
            by = c("ct10" = "ct10"))

census_shapeABC <- census_shapeAB %>%
  left_join(., 
            pitestC,
            by = c("ct10" = "ct10"))


census_shapeABCD <- census_shapeABC %>%
  left_join(., 
            pitestD,
            by = c("ct10" = "ct10"))
  as_tibble()  #The as_tibble() is function that coerce lists and matrices to data frames.

pick<-census_shapeABCD %>%
  dplyr::select(ct10, A_area, B_area, C_area, D_area)

census_shape_joined <- census_shape %>%
  left_join(., 
            pick,
            by = c("ct10" = "ct10"))%>%
  dplyr::mutate(tract_area=st_area(.))

plot(census_shape_joined$geometry)
```
```

% HOLC in tract
```{r}
census_shape_joined_percent <- census_shape_joined %>%
  mutate(percent_area_A=(A_area/tract_area)*100)%>%
  mutate(percent_area_B=(B_area/tract_area)*100)%>%
  mutate(percent_area_C=(C_area/tract_area)*100)%>%
  mutate(percent_area_D=(D_area/tract_area)*100)

```

add in temperature
```{r}
census_shape_joined_percent$ct10 <- as.integer(census_shape_joined_percent$ct10)
LST_per_census_within_HOLC$ID <- as.integer(LST_per_census_within_HOLC$ID )

census_shape_joined_percent_temp <- census_shape_joined_percent %>%
  left_join(., 
            LST_per_census_within_HOLC,
            by = c("ct10" = "ID"))
```

绘制温度和不同等级的评级区域面积比例的关系散点图
```{r}
library(ggplot2)

A <- census_shape_joined_percent_temp %>%
  st_drop_geometry(.)%>%
  #drop_na(.)%>%
  as_tibble(.)%>%
  ggplot(aes(x = as.numeric(percent_area_A) , 
           y = as.numeric(temp)), 
           data=.)+
  geom_point()+
  stat_smooth(method="lm", se=FALSE, size=1)

B <- census_shape_joined_percent_temp %>%
  st_drop_geometry(.)%>%
  #drop_na(.)%>%
  as_tibble(.)%>%
  ggplot(aes(x = as.numeric(percent_area_B) , 
           y = as.numeric(temp)), 
           data=.)+
  geom_point()

C <- census_shape_joined_percent_temp %>%
  st_drop_geometry(.)%>%
  #drop_na(.)%>%
  as_tibble(.)%>%
  ggplot(aes(x = as.numeric(percent_area_C) , 
           y = as.numeric(temp)), 
           data=.)+
  geom_point()


D <- census_shape_joined_percent_temp %>%
  st_drop_geometry(.)%>%
  #drop_na(.)%>%
  as_tibble(.)%>%
  ggplot(aes(x = as.numeric(percent_area_D) , 
           y = as.numeric(temp)), 
           data=.)+
  geom_point()
  


library(cowplot)
all <- plot_grid(
  A, B, C, D,
  labels = "AUTO", ncol = 2
)
all
```



待补统计汇总（R方，p value，t等）



栅格数据聚合到每个tract，可视化map
